Tail Indices and Scale Parameters in Financial Time Series

### 2017-03-06 papers/FX/draft.tex
A revision to the lemma. Applied the lemma to t-distribution.
Added the section about t-distributed equity return.

Routledge and Zin \cite{routledge2010generalized} assumed
a power-law utility function of $C$:
\begin{equation}
  \label{eq:power_utility}
  u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%  
\end{equation}
where $\gamma > 0$. For our discussion it suffices that $u(\cdot)$ is
an increasing function. For simplicity we assume the investor
initially has one unit of wealth. Let $r > 0$ be the return on the
risk free state bond and $X$ be the return on equity. Portfolio weights
are $\left(1-\phi\right)$ and $\phi$ respectively. Then
\begin{equation}
  \label{eq:xxie1}
  %% u(C)=\frac{1}{1-\gamma}V^{1-\gamma} \max \left\{
  %% 0, 1 + (1-\phi) r  +\phi X
  %% \right\}^{1-\gamma} \label{3}%
  C = (1 - \phi) e^r + \phi e^X
\end{equation}
For the distribution of $X$, call it $F(\cdot)$, we assume
%% \subsubsection[in case gamma > 1]{in case $\gamma > 1$}
% \subsection{power-law utility function}
% \subsubsection[in case gamma > 1]{in case $\gamma > 1$}
% In this section we consider power utility functions of
% \eqref{eq:power_utility} with $\gamma > 1$. For convenience,
% let $\xi = \gamma - 1$. In this case, we assume the following
% distribution function $F(\cdot)$ of the equity return:
\begin{equation}
  \label{eq:X_distr}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation}
Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
Equation \eqref{eq:xxie0} can be re-arranged as
\begin{eqnarray}
u(v) &=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) \P[C < \delta v] \nonumber \\
&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q) \label{eq:functional}
\end{eqnarray}
where
\begin{eqnarray*}
  q(r, \phi) &:=& \ln \left( {
    \delta v - (1 - \phi) e^r
    \over
    \phi
} \right) = \ln\left(
e^r + {\delta v - e^r \over \phi}
\right)
\end{eqnarray*}
Note $C < \delta v$ implies
\[
X < q(r, \phi)
\]
Naturally, if an agent invests in a risky asset instead of a riskless
bond, he expects to obtain a higher return from the risky asset than
he is guarranteed from the riskless bond. In our notation,
this means $\delta v > e^r$, which implies $q(r, \phi) > 0$.
The right-hand-side of \eqref{eq:functional} is a functional of the
distribution function of $X$ via \eqref{eq:xxie1}. For convenience
we denote it $\mathcal U(\cdot)$, i.e.
\[
\mathcal U F (\alpha, \beta, K, K', \phi)
= 
\E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q(r, \phi))
\]
Hereafter we abbreviate $q(r, \phi)$ to just $q$. We have
\begin{eqnarray}
  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
  &=&
  \alpha K^\alpha  p
  \int_{-\infty}^0
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \over (K - x)^{\alpha + 1}} dx
  \nonumber \\
  &&
  \beta K'^\beta (1 - p)
  \int_{0}^\infty
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \1{x < q} \over (K' + x)^{\beta + 1}} dx \nonumber \\
  &&
  - b u(\delta v) \left[
    1 - (1 - p) {
      K'^\beta
      \over
      (K' + q)^\beta
    }
  \right]
  \label{eq:xxie1.0}
\end{eqnarray}
In theorem \ref{thrm:I} we show
\[
\max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
\]
is monotonically increasing with $\alpha$ and monotonically decreasing
with $K$. But before that it is worthwhile to state the following
simple observation:
\begin{lemma} \label{lemma:I}
  Assume distribution function $F(x; \theta)$ parameterized by
  $\theta$ is defined on $x \in (a, b)$, $a, b \in \mathbb R \cup
  \{-\infty, \infty\}$, and in addition $F(x; \theta)$ has density
  function $f(x; \theta)$. Assume function $h(\cdot)$ is defined on
  $(a, b)$ and $X \sim F$.
  \begin{enumerate}
  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$, 
    \[
    \pd{\E h(X)}{\theta} > 0
    \]
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$,
    \[
    \pd{\E h(X)}{\theta} > 0
    \]
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$
    \[
    \pd{\E h(X)}{\theta} < 0
    \]
  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that 
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
    \[
    \pd{\E h(X)}{\theta} < 0
    \]
  \end{enumerate}
\end{lemma}
\begin{proof}
  Clearly
  \begin{eqnarray*}
    && \pd{\E h(X)}{\theta} \\
    &=& \int_a^b h(x) \pd{f}{\theta}(x, \theta) dx \\
    &=& \underbrace{\int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx}_{I_1}
    + \underbrace{\int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx}_{I_2}
  \end{eqnarray*}
  \begin{enumerate}
  \item When $h(x)$ is decreasing on $(a, b)$ and $\pd{f}{\theta}(x,
    \theta) > 0$ on $(a, x_0)$
    \[
    I_1 > h(x_0) \int_a^{x_0} \pd{f}{\theta}(x, \theta) dx
    \]
    Similarly, because $\pd{f}{\theta}(x, \theta) < 0$ for $x \in (x_0, b)$ and
    $h(x)$ is decreasing, we have
    \begin{eqnarray*}
      I_2 &=& \int_{x_0}^b -h(x_0)
      \left|\pd{f}{\theta}(x, \theta) \right| dx \\
      &>& -h(x_0)
      \int_{x_0}^b \left| 
        \pd{f}{\theta}(x, \theta)
      \right| dx
    \end{eqnarray*}
    Finally we have
    \begin{eqnarray*}
      && \pd{\E h(X)}{\theta} \\
      &=& I_1 + I_2 \\
      &>& h(x_0) \int_a^b \pd{f}{\theta}(x, \theta) dx \\
      &=& h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx \\
      &=& h(x_0) \opd{\theta} 1 = 0
    \end{eqnarray*}
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
    \begin{eqnarray*}
      I_1 &=&
      \int_a^{x_0} -h(x)
      \left| \pd{f}{\theta}(x, \theta) \right| dx \\
      &>&
      -h(x_0) \int_a^{x_0}
      \left| \pd{f}{\theta}(x, \theta) \right| dx
    \end{eqnarray*}
    and obviously
    \[
    I_2 > h(x_0) \int_{x_0}^b
    \pd{f}{\theta}(x, \theta) dx
    \]
    So
    \[
    \pd{\E h(X)}{\theta}
    = I_1 + I_2
    > h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
    \]
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$

@@

    In this case
    \begin{eqnarray*}
      I_1 &=& \int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx \\
      &<& -h(x_0) \int_a^{x_0} \left| \pd{f}{\theta}(x, \theta) \right| dx
    \end{eqnarray*}
    And
    \begin{eqnarray*}
      I_2 &=& \int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx \\
      &<& h(x_0) \int_{x_0}^b \pd{f}{\theta}(x, \theta) dx
    \end{eqnarray*}
    So
    \[
    \pd{\E h(X)}{\theta}
    = I_1 + I_2
    < h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
    \]
  \end{enumerate}
\end{proof}
With lemma \ref{lemma:I} it is straightforward to show the
monotonicity claimed earlier.
Let
\[
G(\alpha, \beta, K, K') =
\max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
\]
\begin{theorem}
  \label{thrm:I}
  \[
  \pd{G}{\alpha} (\alpha, \beta, K, K') > 0,
  \quad
  \pd{G}{K} (\alpha, \beta, K, K') < 0
  \]
\end{theorem}

\begin{proof}
  Let
  \begin{equation}
    \label{eq:xxie2}
    \hat \phi := \argmax_{0 < \phi \leq 1}
    \mathcal U F(\alpha, \beta, K, K', \phi)
  \end{equation}
  We have
  \[
  G(\alpha, \beta, K, K') = \mathcal U F(\alpha, \beta, K, K', \hat \phi)
  \]
  It follows
  \begin{eqnarray}
    \pd{G}{\alpha}
    &=&
    \pd{\mathcal U F}{\alpha}(\alpha, \beta, K, K', \hat \phi)
    + \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi)
    \pd{\hat \phi}{\alpha}(\alpha, \beta, K, K')
    \label{eq:xxie3}
  \end{eqnarray}
  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
  \begin{equation}
    \label{eq:xxie4}
    \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi) = 0
  \end{equation}
  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
  \begin{eqnarray*}
    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi)
    &>& 0
  \end{eqnarray*}
  From \eqref{eq:xxie1.0}, it follows
  \begin{eqnarray*}
    && \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat\phi)
    \\
    &=& p \opd{\alpha}\E[u((1 - \phi) e^r + \phi e^X) | X < 0]
  \end{eqnarray*}
  The function $u((1 - \phi) e^r + \phi e^X)$ is obviously increasing
  with $X$. Furthermore
  \begin{eqnarray*}
    && \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
    &=&
    - {K^\alpha \over (K - x)^{\alpha + 1}}
    \left[
      \alpha
      \ln\left(
        1 - {x \over K}
      \right) - 1
    \right]
  \end{eqnarray*}
  It is easily checked that when $x < K(1 - e^{1/\alpha}) < 0$,
  \[
  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  \]
  and when $K(1 - e^{1/\alpha}) < x < 0$,
  \[
  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  \]
  This is the second case of lemma \ref{lemma:I}. So we have
  \begin{eqnarray*}
    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) &>& 0
  \end{eqnarray*}
  As for $\pd{G(\alpha, \beta, K, K')}{K}$, by the same argument as for
  $\pd{G(\alpha, \beta, K, K')}{\alpha}$, it suffices to show
  \[
  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi) < 0
  \]
  We have
  \[
  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi)
  = p \opd{K} \E \left[
    u((1 - \phi) e^r + \phi e^X)
    | X < 0
  \right] 
  \]
  and
  \begin{eqnarray*}
    && \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
    &=&
    -\alpha K^{\alpha - 1} {
      \alpha x + K
      \over
      (K - x)^{\alpha + 2}
    }
  \end{eqnarray*}
  Clearly, when $x < -\alpha/K$
  \[
  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  \]
  and when $-\alpha/K < x < 0$
  \[
  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  \]
  So by the third case of lemma \ref{lemma:I} we conclude
  \[
  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi) < 0  
  \]
\end{proof}
Since $u(v)$ given in \eqref{eq:xxie1.0} as a
functional of the distribution function of $X$ is increasing with the
lower tail index $\alpha$ and decreasing with $K$, there is a curve of
equal preference on the plane of $(\alpha, K)$. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e. the estimated values of $\alpha$ and $K$
should appear positively correlated. This is indeed the case in some
real stock return data, e.g. the ``Energy'' and ``Information Technology''
sectors of S\&P 500 companies. This is illustrated in figures
\ref{fig:Energy_alpha_K} and \ref{fig:Information_Technology_alpha_K}.
The confidence bands of the Hill estimates of $\alpha$ of these
sectors are shown in figures \ref{fig:Energy_lower},
\ref{fig:Consumer_Staples_lower} and \ref{fig:Information_Technology_lower}.
% all agents guided by this class of utility
% functions will seek the largest $\alpha$ and the smallest $K$ found in
% a stock group e.g. a sector.

% As a result, one can expect all the
% actively traded equities in the group to feature the same $\alpha$ and
% $K$. However, the estimated value of $K^\alpha$ may vary
% significantly. This can be understood as a result of the smallness of
% $K$, which as revealed in section \ref{sec:hybrid_estimation} is of the
% order of $10^{-2}$, and the inaccuracy of the estimated value of
% $\alpha$, which typically ranges from 2.5 to 5. Exponentiation hugely
% magnifies the inaccuracy in $\alpha$.

>>>>

Routledge and Zin \cite{routledge2010generalized} assumed
a power-law utility function of $C$:
\begin{equation}
  \label{eq:power_utility}
  u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%  
\end{equation}
where $\gamma > 0$. For our discussion it suffices that $u(\cdot)$ is
increasing and concave. Before proceeding any further, it is worth
stating the following simple observation:
\begin{lemma} \label{lemma:I}
  Assume distribution function $F(x; \theta)$ parameterized by
  $\theta$ is defined on $x \in (a, b)$, $a, b \in \mathbb R \cup
  \{-\infty, \infty\}$, and in addition $F(x; \theta)$ has density
  function $f(x; \theta)$. Assume function $h(\cdot)$ is defined on
  $(a, b)$, $X \sim F$, $\E h(X) < \infty$ and $\forall x \in (a, b)$
  \[
  \left| \pd{f(x, \theta)}{\theta} \right| < \infty
  \]
  \begin{enumerate}
  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$, then
    \[
    \pd{\E h(X)}{\theta} > 0
    \]
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$, then
    \[
    \pd{\E h(X)}{\theta} > 0
    \]
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$, then
    \[
    \pd{\E h(X)}{\theta} < 0
    \]
  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that 
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$, then
    \[
    \pd{\E h(X)}{\theta} < 0
    \]
  \end{enumerate}
\end{lemma}
\begin{proof}
  Clearly
  \begin{eqnarray*}
    && \pd{\E h(X)}{\theta} \\
    &=& \int_a^b h(x) \pd{f}{\theta}(x, \theta) dx \\
    &=& \underbrace{\int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx}_{I_1}
    + \underbrace{\int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx}_{I_2}
  \end{eqnarray*}
  \begin{enumerate}
  \item When $h(x)$ is decreasing on $(a, b)$ and $\pd{f}{\theta}(x,
    \theta) > 0$ on $(a, x_0)$
    \[
    I_1 > h(x_0) \int_a^{x_0} \pd{f}{\theta}(x, \theta) dx
    \]
    Similarly, because $\pd{f}{\theta}(x, \theta) < 0$ for $x \in (x_0, b)$ and
    $h(x)$ is decreasing, we have
    \begin{eqnarray*}
      I_2 &=& \int_{x_0}^b -h(x_0)
      \left|\pd{f}{\theta}(x, \theta) \right| dx \\
      &>& -h(x_0)
      \int_{x_0}^b \left| 
        \pd{f}{\theta}(x, \theta)
      \right| dx
    \end{eqnarray*}
    Finally we have
    \begin{eqnarray*}
      && \pd{\E h(X)}{\theta} \\
      &=& I_1 + I_2 \\
      &>& h(x_0) \int_a^b \pd{f}{\theta}(x, \theta) dx \\
      &=& h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx \\
      &=& h(x_0) \opd{\theta} 1 = 0
    \end{eqnarray*}
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
    \begin{eqnarray*}
      I_1 &=&
      \int_a^{x_0} -h(x)
      \left| \pd{f}{\theta}(x, \theta) \right| dx \\
      &>&
      -h(x_0) \int_a^{x_0}
      \left| \pd{f}{\theta}(x, \theta) \right| dx
    \end{eqnarray*}
    and obviously
    \[
    I_2 > h(x_0) \int_{x_0}^b
    \pd{f}{\theta}(x, \theta) dx
    \]
    So
    \[
    \pd{\E h(X)}{\theta}
    = I_1 + I_2
    > h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
    \]
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$

@@

    In this case
    \begin{eqnarray*}
      I_1 &=& \int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx \\
      &<& -h(x_0) \int_a^{x_0} \left| \pd{f}{\theta}(x, \theta) \right| dx
    \end{eqnarray*}
    And
    \begin{eqnarray*}
      I_2 &=& \int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx \\
      &<& h(x_0) \int_{x_0}^b \pd{f}{\theta}(x, \theta) dx
    \end{eqnarray*}
    So
    \[
    \pd{\E h(X)}{\theta}
    = I_1 + I_2
    < h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
    \]
  \end{enumerate}
\end{proof}

Now for simplicity we assume the investor initially has one unit of
wealth. Let $r > 0$ be the return on the risk free state bond and $X$
be the return on equity. Portfolio weights are $\left(1-\phi\right)$
and $\phi$ respectively. Then
\begin{equation}
  \label{eq:xxie1}
  C = (1 - \phi) e^r + \phi e^X
\end{equation}
Equation \eqref{eq:xxie0} can be re-arranged as
\begin{eqnarray}
u(v) &=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) \P[C < \delta v] \nonumber \\
&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q) \label{eq:functional}
\end{eqnarray}
where
\begin{eqnarray*}
  q(r, \phi) &:=& \ln \left( {
    \delta v - (1 - \phi) e^r
    \over
    \phi
} \right) = \ln\left(
e^r + {\delta v - e^r \over \phi}
\right)
\end{eqnarray*}
Note $C < \delta v$ implies
\[
X < q(r, \phi)
\]
Hereafter we abbreviate $q(r, \phi)$ to just $q$. 
Naturally, if an agent invests in a risky asset instead of a riskless
bond, he expects to obtain a higher return from the risky asset than
he is guarranteed from the riskless bond. In our notation,
this means $\delta v > e^r$, which implies $q(r, \phi) > r > 0$.
The right-hand-side of \eqref{eq:functional} is a functional of the
distribution function of $X$ via \eqref{eq:xxie1}. For convenience
we denote it $\mathcal U(\cdot)$, i.e.
\[
\mathcal U F (\alpha, \beta, K, K', \phi)
= 
\E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q(r, \phi))
\]
In the rest of this section, we use $F(\cdot)$ to denote the
distribution function of $X$ and use $f(\cdot)$ to denote its density
function.

\subsection{Pareto-distributed equity return}
In this section, we look into the situation when $X$ has a shifted
Pareto distribution. Assume
\begin{equation}
  \label{eq:X_distr}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation}
Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
We have
\begin{eqnarray}
  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
  &=&
  \alpha K^\alpha  p
  \int_{-\infty}^0
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \over (K - x)^{\alpha + 1}} dx
  \nonumber \\
  &&
  \beta K'^\beta (1 - p)
  \int_{0}^\infty
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \1{x < q} \over (K' + x)^{\beta + 1}} dx \nonumber \\
  &&
  - b u(\delta v) \left[
    1 - (1 - p) {
      K'^\beta
      \over
      (K' + q)^\beta
    }
  \right]
  \label{eq:xxie1.0}
\end{eqnarray}
% In theorem \ref{thrm:I} we show
% \[
% \max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
% \]
% is monotonically increasing with $\alpha$ and monotonically decreasing
% with $K$. But before that it is worthwhile to state the following
% simple observation:
% With lemma \ref{lemma:I} it is straightforward to show the
% monotonicity claimed earlier.
Let
\[
G(\alpha, \beta, K, K') =
\max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
\]
\begin{theorem}
  \label{thrm:I}
  \[
  \pd{G}{\alpha} (\alpha, \beta, K, K') > 0,
  \quad
  \pd{G}{K} (\alpha, \beta, K, K') < 0
  \]
\end{theorem}

\begin{proof}
  Let
  \begin{equation}
    \label{eq:xxie2}
    \hat \phi := \argmax_{0 < \phi \leq 1}
    \mathcal U F(\alpha, \beta, K, K', \phi)
  \end{equation}
  We have
  \[
  G(\alpha, \beta, K, K') = \mathcal U F(\alpha, \beta, K, K', \hat \phi)
  \]
  It follows
  \begin{eqnarray}
    \pd{G}{\alpha}
    &=&
    \pd{\mathcal U F}{\alpha}(\alpha, \beta, K, K', \hat \phi)
    + \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi)
    \pd{\hat \phi}{\alpha}(\alpha, \beta, K, K')
    \label{eq:xxie3}
  \end{eqnarray}
  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
  \begin{equation}
    \label{eq:xxie4}
    \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi) = 0
  \end{equation}
  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
  \begin{eqnarray*}
    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi)
    &>& 0
  \end{eqnarray*}
  From \eqref{eq:xxie1.0}, it follows
  \begin{eqnarray*}
    && \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat\phi)
    \\
    &=& p \opd{\alpha}\E[u((1 - \phi) e^r + \phi e^X) | X < 0]
  \end{eqnarray*}
  The function $u((1 - \phi) e^r + \phi e^X)$ is obviously increasing
  with $X$. Furthermore, the density function of $X | X < 0$ is ${1
    \over p} f(x; \alpha, K)$. It follows
  \begin{eqnarray*}
    && \pd{f(x; \alpha, K)}{\alpha} \\
    &=& \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
    &=&
    - {K^\alpha \over (K - x)^{\alpha + 1}}
    \left[
      \alpha
      \ln\left(
        1 - {x \over K}
      \right) - 1
    \right]
  \end{eqnarray*}
  It is easily checked that when $x < K(1 - e^{1/\alpha}) < 0$,
  \[
  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  \]
  and when $K(1 - e^{1/\alpha}) < x < 0$,
  \[
  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  \]
  This is the second case of lemma \ref{lemma:I}. So we have
  \begin{eqnarray*}
    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) &>& 0
  \end{eqnarray*}
  As for $\pd{G(\alpha, \beta, K, K')}{K}$, by the same argument as for
  $\pd{G(\alpha, \beta, K, K')}{\alpha}$, it suffices to show
  \[
  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi) < 0
  \]
  We have
  \[
  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi)
  = p \opd{K} \E \left[
    u((1 - \phi) e^r + \phi e^X)
    | X < 0
  \right] 
  \]
  and
  \begin{eqnarray*}
    && \pd{f(x; \alpha, K)}{K} \\
    &=& \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
    &=&
    -\alpha K^{\alpha - 1} {
      \alpha x + K
      \over
      (K - x)^{\alpha + 2}
    }
  \end{eqnarray*}
  Clearly, when $x < -\alpha/K$
  \[
  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  \]
  and when $-\alpha/K < x < 0$
  \[
  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  \]
  So by the third case of lemma \ref{lemma:I} we conclude
  \[
  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi) < 0  
  \]
\end{proof}
Since $u(v)$ given in \eqref{eq:xxie1.0} as a
functional of the distribution function of $X$ is increasing with the
lower tail index $\alpha$ and decreasing with $K$, there is a curve of
equal preference on the plane of $(\alpha, K)$. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e. the estimated values of $\alpha$ and $K$
should appear positively correlated. This is indeed the case in some
real stock return data, e.g. the ``Energy'' and ``Information Technology''
sectors of S\&P 500 companies. This is illustrated in figures
\ref{fig:Energy_alpha_K} and \ref{fig:Information_Technology_alpha_K}.
The confidence bands of the Hill estimates of $\alpha$ of these
sectors are shown in figures \ref{fig:Energy_lower},
\ref{fig:Consumer_Staples_lower} and
\ref{fig:Information_Technology_lower}.

\subsection{Student's t-distributed equity return}
It is a common practice to use Student's t-distribution to model the
stationary distribution of equity returns. So it is of interest to
find out what implications this distribution has when it is combined
with the PDA preference. Formally we assume
\[
f(x; \nu) = c(\nu) \left(
  1 + {x^2 \over \nu}
\right)^{-(\nu + 1)/2}
\]
where $\nu > 1$ and
\[
c(\nu) = {
  \Gamma({\nu + 1 \over 2})
  \over
  \Gamma(\nu/2) \sqrt{\nu \pi}
}
\]
Then it follows
\begin{eqnarray*}
\pd{f(x, \nu)}{\nu} &=& {
  c(\nu) x^2 (\nu + 1) + (2 \nu x^2 + 2 \nu^2) c'(\nu)
  -
  \nu c(\nu) (x^2 + \nu) \ln(1 + x^2/\nu)
  \over
  2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
}
\end{eqnarray*}
The nummerator of the right side of $\pd{f(x, \nu)}{\nu}$, if seen as
a function of $x^2$, has a single root:
\begin{equation}
  \label{eq:xxie5}
  x^2 = \nu\exp\left\{
    W\left[
      -\left(1 + {1 \over \nu}\right)
      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
    \right]
    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
  \right\} - \nu
\end{equation}
where $W(\cdot)$ is the principle branch of the Lambert $W$
function. and $c'(\cdot)$ is the derivative of $c(\cdot)$. To check
the right side of \eqref{eq:xxie5} for positivity, we first note
$c'(\nu) > 0$:
\[
c'(\nu) = {
  \pi \Gamma(\nu/2 + 1/2) \left[
    \nu \Psi(\nu/2 + 1/2) - \nu \Psi(\nu/2) - 1
  \right]
  \over
  2 \Gamma(\nu/2) (\pi \nu)^{3/2}
}
\]
where $\Psi(\cdot)$ is the digamma function:
\[
\Psi(x) = \td{\ln[\Gamma(x)]}{x}
\]
Since for all $x > 0$, $\Psi(x) > 0$ and $\nu > 1$, it is clear
$c'(\nu) > 0$. Furthermore, we recall
$W(\cdot)$ is increasing on its principle branch. So
\begin{eqnarray*}
  &&
  W\left[
    -\left( 1 + {1 \over \nu} \right)
    e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
  \right]
  + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
  &>& 
  W\left[
    -\left( 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \right)
    e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
  \right]
  + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
  &=& W(-y e^{-y}) + y
\end{eqnarray*}
where
\[
y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
\]
Now notice
\[
\ln(y e^{-y}) = \ln(y) - y
\]
is a decreasing function. Thus $-y e^{-y}$ is an increasing
function. Hence we have
\begin{eqnarray*}
  W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
\end{eqnarray*}
Now it is clear
\[
\nu\exp\left\{
    W\left[
      -\left(1 + {1 \over \nu}\right)
      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
    \right]
    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
  \right\} - \nu > 0
\]
